---
title: "ASSIGNMENT1"
author: "Emine Uysal"
date: "2024-07-10"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##                                          CIND 123
##                                Data Analytics: Basic Methods
###                                      Assignment 1 (10%)


[Student number:501304049]
                                      
####                                      INSTRUCTIONS

This assignment can be submitted using either Python or R, whichever you prefer.
If using R, you must submit an RMD file with its knitted file (PDF or HTML). To learn more about knitting and R markdown, visit R Markdown.
If using Python, you must submit an IPYNB file and its exported PDF/HTML with clearly printed/shown answers.
Failing to submit both files ({RMD + knitted PDF/HTML} OR {IPYNB + PDF/HTML}) will be subject to a 30% mark deduction.

NOTE: IF YOU USE R STUDIO , YOU SHOULD NEVER HAVE install.packages IN YOUR CODE; OTHERWISE, THE Knit OPTION WILL RAISE AN ERROR.
COMMENT OUT ALL PACKAGE INSTALLATIONS BUT KEEP library() CALLS.

NOTE: If you answer the questions in R, all your answers should be in R (ignore Python questions). If you answer the questions in Python, all your answers should be
in Python (ignore R questions). You are not allowed to switch languages in this assignment.

#Question 1 (50 points)

##Q1a (15 points)

Create and print a vector x (or list in Python) with all odd numbers from 5 to 120 and a vector y containing multiples of 5 in the same range.

HINT: use seq() in R or range() in Python.

```{r}
x <- seq(5, 120, by = 2)
print(x)
y <- seq(5, 120, by = 5)
print(y)
```

Calculate the sum of the lengths of the vectors x and y .

HINT: use length() in R or len() in Python.

```{r}
length_x <- length(x)
length_y <- length(y)
sum_length <- length_x + length_y
print(sum_length)
```
##Q1b (15 points)

Create a new vector (or list in Python), y_cube , with the cube of elements at indices 2, 4, 6, 10, 15, 18, 21, and 25 from the variable y .

HINT: Use indexing rather than a for loop.

```{r}
# Example vector y (ensure it has at least 25 elements)
y <- 1:30
# Specified indices
indices <- c(2, 4, 6, 10, 15, 18, 21, 25)
# Create y_cube by cubing the elements at the specified indices
y_cube <- y[indices]^3
# Print y_cube to see the result
print(y_cube)

```
Calculate the mean and median of the LAST five values from y_cube .

```{r}
mean_last_five <- mean(y_cube[(length(y_cube)-4):length(y_cube)])
print(mean_last_five)
median_last_five <- median(y_cube[(length(y_cube)-4):length(y_cube)])
print(median_last_five)
```
#Q1c (10 points)

For a given factor variable
factorVar <- factor(c("2.5", "7.1", "6.3", "4.8")) in R or
factorVar = ["2.5", "7.1", "6.3", "4.8"] in Python ,
Would it be correct to use the following commands to convert factor to number?
as.numeric(factorVar) in R or map(float, factorVar) in Python.
If not, explain your answer and provide the correct one.

```{r}
###Answer: No, it would not be correct to use as.numeric(factorVar) in R or map(float, factorVar) in Python directly.

###Explanation: In R, if you use as.numeric() on a factor variable, it won't provide the numeric values of the levels. Instead, it will return the internal integer codes R uses to denote the factor levels.To convert a factor to numeric correctly, you should first convert it to character and then to numeric in R. Like that:

factorVar <- factor(c("2.5", "7.1", "6.3", "4.8"))
numericVar <- as.numeric(as.character(factorVar))
```
##Q1d (10 points)

A comma-separated values file dataset.csv consists of missing values represented by Not A Number (null) and question mark (?). Read this type of file in R or Python and preprocess the missing values.

NOTE: Please ensure you have saved the dataset.csv file in your current working directory.

```{r}
data <- read.csv("C:/Users/emine/OneDrive/Masaüstü/ASSIGNMENT1-CIND123/dataset.csv", na.strings = c("null", "?"))
for(i in which(sapply(data, is.numeric))){
  data[is.na(data[,i]), i] <- mean(data[,i], na.rm = TRUE)
}
```

Print the head of the file to ensure everything is correct. DO NOT print the whole file.

```{r}
head(data)
```
#Question 2 (50 points)

##Q2a (15 points)

Compute: $$\sum_{n=4}^{25} \frac{(−1)^{n+1}}{(2n+1)!}$$
HINT: Use factorial(n) in R or math.factorial(n) in Python to compute .
  

```{r}
# Load the factorial function
library(gmp) # for large factorials
```

```{r}
# Define the sum
sum_result <- 0
```
```{r}
# Compute the sum
for (n in 4:25) {
  term <- ((-1)^(n+1)) / factorial(2*n + 1)
  sum_result <- sum_result + term
}
```
```{r}
# Print the result
print(sum_result)
```
##Q2b (15 points)
  

Compute: 	$$\prod_{x = 2}^{6} (3n+\frac{2}{3^n})$$

  
NOTE: The symbol represents multiplication.
  
```{r}
# Initialize the product result
product_result <- 1
```
```{r}
# Loop through n from 2 to 6
for (n in 2:6) {
    term <- 3 * n + 2 / (3^n)
    product_result <- product_result * term
  }
```
```{r}
# Print the result
print(product_result)
```
Q2a and Q2b Guidance

For Q2a and Q2b, you'll need to apply basic mathematical concepts in R or Python. If you understand how to code in R or Python and the basic math operations (like
factorial, powers, sums, and products), you're all set!
Example for Understanding: Let's start with a simpler example: Compute the sum of the first 5 square numbers. In mathematical terms, this is:
which translates to .

  In R, you could solve this by:
  
numbers <- 1:5 # Create a vector of numbers 1 to 5
squared_numbers <- numbers^2 # Square each element of the vector
sum_squares <- sum(squared_numbers) # Sum up the squared numbers
print(sum_squares)

  In Python, you could solve this by:
  
numbers = range(1, 6) # Create a list of numbers from 1 to 5
squared_numbers = [n**2 for n in numbers] # Square each element of the list
sum_squares = sum(squared_numbers) # Sum up the squared numbers
print(sum_squares)

Now, come back to Q2a and Q2b. Your task in those questions is similar. You'll use loops to iterate over a range of numbers, applying the specified mathematical operations at each step. Remember to use built-in functions like factorial() , sum() , and prod() in R, or math.factorial() , sum() , and math.prod() in Python when needed.


#Q2c (10 points)

Describe what the following R command does: rep(1:3, each = 2)
NOTE: If you answer the questions in Python, provide an equivalent command and explain its behavior.

```{r}
###Answer: The R command rep(1:3, each = 2) generates a vector that repeats each of the numbers from 1 to 3 twice.

###Explanation: The rep() function in R is used to replicate the values in x. The each argument specifies how many times each element should be repeated. So, rep(1:3, each = 2) means "repeat each of the numbers from 1 to 3 twice". The output of this command would be a vector: 1, 1, 2, 2, 3, 3.
```

#Q2d (10 points)

Describe the purpose of is.logical() , is.character() , is.numeric() , and is.na() functions in R.
In Python, describe the purpose of isinstance() , type() , and pd.isna() functions from the pandas library.
Please use x <- c("a", FALSE, "b", NA, 2, TRUE) in R or x = ["a", False, "b", None, 2, True] in Python to explain your description

```{r}
x <- c("a", FALSE, "b", NA, 2, TRUE)

is.logical(x)   # Checks if all elements are logical
is.character(x) # Checks if all elements are character strings
is.numeric(x)   # Checks if all elements are numeric
is.na(x)        # Checks for NA values in the vector
```

